Precise spiking motifs in neurobiological and neuromorphic data

### 2022-12-19 manuscript.tex
Update on Overleaf.
in sync: spade

%Let us now focus on distances defined based on a generative model formed by using precise spiking motifs. 
Concerning spike timings,~\citet{levakova_review_2015} \change[Antoine]{has}{have} \change[JN]{reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning}{reviewed existing methods for estimating the latency of neural responses that include Bayesian binning.}. \change[JN]{More broadly }{Alternatively }, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}. This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}. A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}. By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}. This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance. The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \add[LP]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.

@@

%----------------------------%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
\caption{\textbf{Detecting motifs using SPADE.} We use the SPADE algorithm~\citep{stella_3d-spade_2019} on the data generated in  Figure~\ref{fig:diesman} to label different precise spike motifs which are denoted by different colors. Spikes are labelled by the same color when they belong to the same motif. As the synfire chain gets more precise, the detected patterns grow in size.}\label{fig:Diesmann_et_al_1999_spade}
\end{figure}
%----------------------------%

>>>>

%Let us now focus on distances defined based on a generative model formed by using precise spiking motifs. 
Concerning spike timings,~\citet{levakova_review_2015} \change[Antoine]{has}{have} \change[JN]{reviewed the methods for neuronal response latency estimation exists and includes Bayesian binning}{reviewed existing methods for estimating the latency of neural responses that include Bayesian binning.}. \change[JN]{More broadly }{Alternatively }, unitary event analysis can be performed by a statistical model of coincidence detection~\citep{grun_unitary_2002-1}. This was extensively used in detecting above chance significant synchronous patterns~\citep{grun_unitary_2010}, in particular in recordings of pairs of neurons (see~\citep{riehle_spike_1997} for instance), and has been extended to non-stationary data~\citep{grun_unitary_2002}. A method to detect significant patterns of synchronous spiking in a subset of massively parallel spike trains in the presence of background activity can be defined using the statistical evaluation of synchronous spike patterns extracted by frequent item set mining~\citep{torre_statistical_2013}. By the same group, the SPADE, CAD or ASSET algorithms are methods for identification of spike patterns in massively parallel spike trains (the spiking activity of tens to hundred(s) of neurons recorded in parallel) by identifying fine temporal correlations in the$~\ms$ precision range~\citep{quaglio_methods_2018}. This was recently extended in~\citep{stella_3d-spade_2019} in order to find re-occurring patterns in parallel spike train data, and to determine their statistical significance. The extension improves the performance in the presence of patterns with different durations, as demonstrated by application to various synthetic data \add[Laurent]{such as the synthetic data for synfire chains (see Figure}~\ref{fig:Diesmann_et_al_1999_spade}), such as surrogates generated to evaluate precisely timed higher-order spike correlations~\citep{stella_comparing_2022}.

@@

%----------------------------%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{figures/Diesmann_et_al_1999_spade.pdf}
\caption{\textbf{Detecting motifs using SPADE.} We used the SPADE algorithm~\citep{stella_3d-spade_2019} by adapting their \href{https://elephant.readthedocs.io/en/latest/tutorials/spade.html}{tutorial} on the data generated in  Figure~\ref{fig:diesman}. This allowed to label different precise spike motifs which are denoted by different colors. Spikes belonging to the same motif have the same color. %As the synfire chain gets more precise, the detected patterns grow in size.
}\label{fig:Diesmann_et_al_1999_spade}
\end{figure}
%----------------------------%
